Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach

### 2024-11-03 paper/paper.qmd
update paper with new model data

@fig-pollstertime shows the trend of percentage support for Kamala
Harris over time, with data points and smoothed trend lines for each
pollster. The x-axis represents the dates from August through October,
while the y-axis indicates the percentage of respondents supporting
Harris. Each color corresponds to a specific polling organization, with
prominent pollsters such as Beacon/Shaw, Ipsos, Siena/NYT, Emerson,
Quinnipiac, and YouGov. The smoothed lines reveal subtle trends over
time, with some pollsters like Emerson and Beacon/Shaw showing a slight
upward trend, while others like Siena/NYT display a downward trend.
@fig-pollstertime highlights the variability in poll results across
different organizations, reflecting each pollster's methodology and
sample.

@@

@fig-numericgradetime illustrates Harris's support percentages over
time, with each data point representing poll results colored by the
numeric grade of the pollster. The smoothing lines for each numeric
grade remain subtle, indicating minor variations in support trends over
time across different pollster quality levels. Most points are clustered
around the 50% support level, suggesting a generally stable voter
sentiment for Harris, with only slight fluctuations across numeric
grades. By limiting the y-axis, outliers and extreme variations are
minimized, resulting in a clearer and more interpretable visualization.
This approach emphasizes the central trend, allowing for clearer
comparisons of support levels across pollsters of varying reliability
without distraction from extreme values.

@@

ggplot(harris_data %>% filter(pct <= 100), aes(x = end_date, y = pct, color = factor(numeric_grade))) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, span = 0.3) +
  labs(y = "Harris Percent", x = "Date", color = "Numeric Grade") +
  scale_color_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```

@@

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations.The model summary of support for Harris is shown in @tbl-summary1, support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 45.10 and 59.28 respectively in the unweighted models,
represent their baseline support levels when all predictors are at
reference values. The high ( R\^2 ) values of 0.738 for Harris’s
unweighted model and 0.860 for her weighted model, along with similar
values of 0.635 and 0.890 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

@@

The results underscore the significance of **national_poll** as key predictors for both candidates' support, highlighting how national trends and pollster differences impact the
reported support levels, as polls from the National level increase the
support of Harris by 3.262%, decrease the support for Trump by 7.943%.
The **population** predictor indicates that the type of respondent group
contributes meaningfully to variations in support. With "likely voters"
as the baseline level, the coefficient for "registered voters"
(populationrv) suggests a lower level of support for both Harris and
Trump in this group compared to likely voters. Specifically, the support
of registered voters is 2.1095 (unweighted model) and 1.0322 (weighted
model) lower than the likely voters. This implies that, on average,
support among registered voters is lower than that among likely voters.

@@

@fig-support-trend illustrate the predicted support for Kamala Harris
and Donald Trump from August 1, 2024, to October 27, 2024, using both
unweighted and weighted linear regression models. In the
@fig-support-trend-1 unweighted model, we see that Trump's predicted
support generally hovers between 45% and 50%, while Harris's support
ranges from 35% to 45%. Both candidates' support levels appear
relatively stable across the polling period, with minor fluctuations.
This model, which treats all polls equally, suggests a consistent lead
for Trump, albeit within a fairly close range, as shown by the
overlapping grey points representing individual poll results.

The weighted model @fig-support-trend-2 refines these predictions by
applying weights based on recency and sample size, which reduces the
influence of older or smaller polls. This approach results in smoother
trend lines for both candidates, with less apparent variability compared
to the unweighted model @fig-support-trend-1. Trump's support remains
around the mid-40% range, while Harris’s support stays in the high 30%
range. The weighting emphasizes more recent and larger polls, likely
providing a more accurate reflection of current public opinion. Overall,
both models consistently indicate a lead for Trump, with the weighted
model slightly stabilizing the predictions.

@@

set.seed(233)
plan(multisession)
####### Prediction for Harris ######
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 2000
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 2000, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 2000, replace = TRUE)
)|>
  mutate(
    national_poll = if_else(state == "National", 1, 0),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Kamala Harris"
  )

####### Prediction for Trump ######
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 2000
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 2000, replace = TRUE), 
                    levels = levels(trump_data$pollster)),
 population= factor(sample(levels(trump_data$population), size = 2000, replace = TRUE), 
                    levels = levels(trump_data$population)),
   state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 2000, replace = TRUE), 
                    levels = levels(trump_data$state)),  
  sample_size =sample(1:2300, size = 2000, replace = TRUE),
  pollster_quality= sample(2:4, size = 2000, replace = TRUE)
)|>
  mutate(
   national_poll = if_else(state == "National", 1, 0),
      recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
   pollster_quality_weight = pollster_quality / 4,  
    recency_weight = exp(-recency * 0.1),  
    combined_weight = recency_weight * sample_size_weight,
    candidate="Donald Trump"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# Predictions for Harris
plan(multisession)
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 5% in Predicted Support in the 2024 US Election (From August 1, 2024, to October 27, 2024)"
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
##weighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+
  labs(x = "End Date of the Poll",
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```

@@

In the @fig-support-state-1 , support for Kamala Harris and Donald Trump
fluctuates widely across battleground states, with Harris holding slight
leads in Pennsylvania and North Carolina but facing tight competition
overall. Without weighting, polls contribute equally, creating more
volatility.

The @fig-support-state-2 stabilizes these trends by emphasizing recent,
larger polls. Harris's support becomes more consistent, particularly in
Pennsylvania, North Carolina, and Nevada, where she shows a clearer
lead. This weighted approach provides a more refined snapshot,
highlighting Harris’s likely advantage in these key states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, North Carolina, and Nevada by Model Prediction"
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)

@@

@fig-support-pollster depict the predicted support for Kamala Harris and
Donald Trump in the 2024 U.S. election, from August 1, 2024, to October
27, 2024, using a weighted multiple linear regression model. The red
line shows fluctuations in Trump’s support, generally between 40% and
55%, with some noticeable peaks and dips, indicating variability over
time. In contrast, Harris’s support, represented by the blue line,
remains more stable, consistently ranging around 40% to 50%, suggesting
less volatility in voter sentiment for her.

Major Pollsters are represented by different colors, with circle sizes
reflecting weights based on factors like recency and sample size. Larger
circles correspond to higher-weighted polls, indicating they had a
greater influence on the predictions. The model incorporates national
poll data, adjustments for pollster effects, and population
considerations, allowing for a refined prediction approach. This
approach ensures that more recent and larger polls play a more
significant role, capturing shifts in voter sentiment as the election
approaches.

@@

ggplot(combined_data_total, aes(x = end_date, y = pct)) +
  geom_point(aes(fill = pollster_group), alpha = 0.7, shape = 21) + 
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_unweighted), color = "blue", size = 1) +
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_unweighted), color = "red", size = 1) +
  labs(
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    color = "Candidate",
    fill = "Top 5 Pollster",
    caption = "Candidate Support Percentage = national_poll + pollster + population + state"
  )+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(brewer.pal(10, "Set1"), top_pollsters))) + 
  theme_classic() +
  theme(
    legend.position = "right",
    legend.box = "vertical",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

ggplot(combined_data_total, aes(x = end_date, y = pct, size = combined_weight)) +
  geom_point(aes(fill = pollster_group), alpha = 0.7, shape = 21) + 
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_weighted), color = "blue", size = 1) +
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_weighted), color = "red", size = 1)+
  labs(
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    size = "Weight",
    color = "Candidate",
    fill = "Top 5 Pollster",
    caption = "Weights: recency_weight * sample_size_weight \nCandidate Support Percentage = national_poll + pollster + population + state"
  )+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(brewer.pal(10, "Set1"), top_pollsters))) +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.box = "vertical",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

>>>>

@fig-pollstertime shows the trend of percentage support for Kamala
Harris over time, with data points and smoothed trend lines for each
pollster. The graph shows the trend of the percentage of respondents supporting
Harris from August through October. Each color corresponds to a specific polling organization, with
prominent pollsters such as Beacon/Shaw, Ipsos, Siena/NYT, Emerson,
Quinnipiac, and YouGov. The smoothed lines reveal subtle trends over
time, with some pollsters like Emerson and Beacon/Shaw showing a slight
upward trend, while others like Siena/NYT display a downward trend.
@fig-pollstertime highlights the variability in poll results across
different organizations, reflecting each pollster's methodology and
sample.

@@

@fig-numericgradetime illustrates Harris's support percentages over
time, with each data point representing poll results colored by the
numeric grade of the pollster. The smoothing lines for each numeric
grade remain subtle, indicating minor variations in support trends over
time across different pollster quality levels. Most points are clustered
around the 50% support level, suggesting a generally stable voter
sentiment for Harris, with only slight fluctuations across numeric
grades. By limiting the maximum value of support at 80%, outliers and extreme variations are minimized, resulting in a clearer and more interpretable visualization.
This approach emphasizes the central trend, allowing for clearer
comparisons of support levels across pollsters of varying reliability
without distraction from extreme values.

@@

ggplot(harris_data %>% filter(pct <= 80), aes(x = end_date, y = pct, color = factor(numeric_grade))) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, span = 0.3) +
  labs(y = "Support for Harris", x = "Date", color = "Numeric Grade") +
  scale_color_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```

@@

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations.The model summary of support for Harris is shown in @tbl-summary1, support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 44.70 and 59.28 respectively in the unweighted models, while in weighted models, it's estimated at 45.2 and 54.8,
represent their baseline support levels when all predictors are at
reference values. The high ( R\^2 ) values of 0.824 for Harris’s
unweighted model and 0.892 for her weighted model, along with similar
values of 0.849 and 0.916 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

@@

The results underscore the significance of **national_poll** as key predictors for both candidates' support, highlighting how national trends and pollster differences impact the
reported support levels, as polls from the National level increase the
support of Harris by 3.47%/4.27% , decrease the support for Trump by 7.61%/7.17%.
The **population** predictor indicates that the type of respondent group
contributes meaningfully to variations in support. With "likely voters"
as the baseline level, the coefficient for "registered voters"
(populationrv) suggests a lower level of support for both Harris and
Trump in this group compared to likely voters. Specifically, the support for Harris
of registered voters is 1.372 (unweighted model) or 0.529 (weighted
model) lower than the likely voters. This implies that, on average,
support among registered voters is lower than that among likely voters.

@@

@fig-support-trend illustrate the predicted support for Kamala Harris
and Donald Trump from August 1, 2024, to October 27, 2024, using both
unweighted and weighted linear regression models. In the
@fig-support-trend-1 unweighted model, we see that Trump's predicted
support generally hovers between 40% and 50%, while Harris's support
ranges from 45% to 50%. Both candidates' support levels appear
relatively stable across the polling period, with minor fluctuations.
This model, which treats all polls equally, suggests a consistent lead for Harris, albeit within a fairly close range.

The weighted model @fig-support-trend-2 refines these predictions by
applying weights based on recency and sample size, which reduces the
influence of older or smaller polls. This approach results in smoother
trend lines for both candidates, with less apparent variability compared
to the unweighted model @fig-support-trend-1. Trump's support remains
around the 40% to 50% range, while Harris’s support stays in about 45% to 52%  range. The weighting emphasizes more recent and larger polls, likely
providing a more accurate reflection of current public opinion. Overall, both models consistently indicate a lead for Harris, with the weighted model slightly stabilizing the predictions. The average support for Harris and Trump are both consist for two models, which predicts that Harris will win on about 49%, leads Trump by about 2%.

@@

set.seed(233)
plan(multisession)
####### Prediction for Harris ######
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 700
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 700, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 700, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 700, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 700, replace = TRUE)
)|>
  mutate(
    national_poll = if_else(state == "National", 1, 0),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Kamala Harris"
  )

####### Prediction for Trump ######
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 700
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 700, replace = TRUE), 
                    levels = levels(trump_data$pollster)),
 population= factor(sample(levels(trump_data$population), size = 700, replace = TRUE), 
                    levels = levels(trump_data$population)),
   state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 700, replace = TRUE), 
                    levels = levels(trump_data$state)),  
  sample_size =sample(1:2300, size = 700, replace = TRUE)
)|>
  mutate(
   national_poll = if_else(state == "National", 1, 0),
      recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),  
    combined_weight = recency_weight * sample_size_weight,
    candidate="Donald Trump"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 5% in Predicted Support in the 2024 US Election (From August 1, 2024, to October 27, 2024)"
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted),size = 0.2)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
##weighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.2)+
  labs(x = "End Date of the Poll",
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```

@@

In the @fig-support-state-1 , support for Kamala Harris and Donald Trump
fluctuates widely across battleground states, with Harris holding significant
leads in Michigan, and Pennsylvania.  Harris also hold advantage in Wisconsin but facing tight competition. Overall, without weighting, polls contribute equally, creating more
volatility.

The @fig-support-state-2 stabilizes these trends by emphasizing recent,
larger polls. Harris's support becomes more consistent, particularly in
Pennsylvania, Michigan, and Wiscosin, where she shows a clearer
lead. This weighted approach provides a more refined snapshot,
highlighting Harris’s likely advantage in these key states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, North Carolina, and Wisconsin by Model Prediction"
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)

@@

@fig-support-pollster depict the predicted support for Kamala Harris and
Donald Trump in the 2024 U.S. election, from August 1, 2024, to October
27, 2024, using the unweighted and weighted multiple linear regression model. The red
line shows average fluctuations in Trump’s support, generally between 46% and
48%, with some noticeable peaks and dips, indicating variability over
time. In contrast,the average Harris’s support, represented by the blue line,
remains more stable, consistently ranging around 47% to 48%, suggesting
less volatility in voter sentiment for her.

Major Pollsters are represented by different colors, with circle sizes
reflecting weights based on factors like recency and sample size. In the weighted model, larger
circles correspond to higher-weighted polls, indicating they had a
greater influence on the predictions. The model incorporates national
poll data, adjustments for pollster effects, and population
considerations, allowing for a refined prediction approach. This
approach ensures that more recent and larger polls play a more
significant role, capturing shifts in voter sentiment as the election
approaches.

@@

ggplot(combined_data_total, aes(x = end_date, y = pct)) +
  geom_point(aes(fill = pollster_group), alpha = 0.7, shape = 21) + 
  geom_smooth(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_unweighted), color = "blue", size = 1) +
  geom_smooth(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_unweighted), color = "red", size = 1) +
  labs(
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    color = "Candidate",
    fill = "Top 5 Pollster",
    caption = "Candidate Support Percentage = national_poll + pollster + population + state"
  )+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(brewer.pal(10, "Set1"), top_pollsters))) + 
  theme_classic() +
  theme(
    legend.position = "right",
    legend.box = "vertical",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

ggplot(combined_data_total, aes(x = end_date, y = pct, size = combined_weight)) +
  geom_point(aes(fill = pollster_group), alpha = 0.7, shape = 21) + 
  geom_smooth(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_weighted), color = "blue", size = 1) +
  geom_smooth(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_weighted), color = "red", size = 1)+
  labs(
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    size = "Weight",
    color = "Candidate",
    fill = "Top 5 Pollster",
    caption = "Weights: recency_weight * sample_size_weight \nCandidate Support Percentage = national_poll + pollster + population + state"
  )+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(brewer.pal(10, "Set1"), top_pollsters))) +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.box = "vertical",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
